\textbf{{\Large {Recursive algorithms, branching coefficients and applications}}}

### 2010-06-15 AffineLieAlgebras/InformalNotes/article.tex
Conformal embeddings rewritten. Text rearranged for the future incorporation of coset models.

\documentclass[a4paper,12pt]{article}
\usepackage{ucs}
\usepackage[unicode,verbose]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{pb-diagram}
\usepackage{multicol}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian]{babel}
\usepackage{cmap}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\pagestyle{plain}

@@

\begin{document}
\thispagestyle{empty}

@@

\section{Examples}
\label{sec:examples}

\subsection{Finite dimensional Lie algebras}
\label{sec:finite-dimens-lie}

\subsubsection{Regular embedding of $A_1$ into $B_2$}
\label{sec:regul-embedd-a_1}

@@

\subsubsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the regular embedding $B_2 \longrightarrow B_4$.
We calculate the branching coefficients for the fundamental vector representation of $B_4$.
The corresponding Dynkin diagrams are in the Figure \ref{fig:dynkin}.
\begin{figure}[h]
  \centering
  \includegraphics[width=50mm]{B4_B2_2A1.pdf}
  \caption{Dynkin diagrams for the regular embedding $B_2\subset B_4$. Note the dropped node and the roots of $\mathfrak{a}_{\bot}$. }
  \label{fig:dynkin}
\end{figure}

@@

\begin{figure}[p]
  \centering
  \includegraphics[height=80mm]{B4_B2_fan2.pdf}
  \caption{The fan for $B_2\subset B_4$}
  \label{fig:B4B2Fan}
\end{figure}
Applying the recurrent relation (\ref{recurrent-relation}) we obtain the
following branching coefficients:
\begin{multline}
  \label{eq:24}
  \pi_{\mathfrak{a}} \left(ch L^{(0,1,0,2)}_{B_4}\right) = 6 \; ch L^{(0,0)}_{B_2}+ 60
  \; ch L_{B_2}^{(0,2)}+ 30 \; ch L_{B_2}^{(1,0)}+ 19 \; ch L_{B_2}^{(2,0)}+\\
  40 \; ch L_{B_2}^{(1,2)}+ 10 \; ch L_{B_2}^{(2,2)}.
\end{multline}

\subsection{Affine into affine embedding}
\label{sec:embedd-affine-algebr}

%\newpage
\section{Applications to the conformal field theory}
\label{sec:phys-appl}

Branching coefficients for an embedding of affine Lie subalgebra into
affine Lie algebra can be used to construct modular invariant
partition functions for Wess-Zumino-Novikov-Witten models of conformal field theory (\cite{difrancesco1997cft}, \cite{Walton:1999xc}, \cite{walton1989conformal}, \cite{schellekens1986conformal}). In these models current algebras are affine Lie algebras.  For the construction to be valid the embedding is required to be conformal, which means that the central charge of the subalgebra is equal to the central charge of the algebra:
\begin{equation}
  \label{eq:31}
  c(\mathfrak{a})=c(\mathfrak{g})
\end{equation}

Let $X^{a_j}_{-n_j}$ and $\tilde{X}^{a'_j}_{-n_j}$ be the lowering generators for $\mathfrak{g}$ and for $\mathfrak{a}\subset\mathfrak{g}$ correspondingly. Let $\pi_{\mathfrak{a}}$ be the projection operator of $\pi_{\mathfrak{a}}:\mathfrak{g}\longrightarrow \mathfrak{a}$. In the theory attributed to $\mathfrak{g}$ with the vacuum $\left|\lambda\right>$ the states can be described as
\begin{equation}
  \label{eq:109}
  X^{a_1}_{-n_1}X^{a_2}_{-n_2}\dots\left|\lambda\right>\quad n_1\geq n_2\geq \dots>0.
\end{equation}
And for the sub-algebra $\mathfrak{a}$ the corresponding states are
\begin{equation}
  \label{eq:110}
  \tilde{X}^{a'_1}_{-n_1}\tilde{X}^{a'_2}_{-n_2}\dots\left|\pi_{\mathfrak{a}}(\lambda)\right>.
\end{equation}
The $\mathfrak{g}$-invariance of the vacuum entails its $\mathfrak{a}$-invariance, but this is not the case for the energy-momentum tensor. So the energy-momentum tensor of the larger theory should consist only of the generators of $\tilde{X}$. Then $T_{\mathfrak{g}}=T_{\mathfrak{a}}\Rightarrow c(\mathfrak{g})=c(\mathfrak{a})$. This leads to the equation
\begin{equation}
  \label{eq:111}
  \frac{k\;\mathrm{dim}\,\mathfrak{g}}{k+g}=\frac{x_e k\; \mathrm{dim}\,\mathfrak{a}}{x_ek+a}
\end{equation}
Here $x_e$ is the embedding index and $g$, $a$ are the dual Coxeter numbers for the  corresponding algebras.

@@

If we have modular-invariant partition function for the fields described by a representation of the algebra $\mathfrak{g}$ this modular invariance is preserved by the projection on the subalgebra $\mathfrak{a}$, but we need also the preservation of the conformal invariance. So we should select only those highest-weight modules of the algebra $\mathfrak{g}$ for which the relation (\ref{eq:111}) holds.

This relation and the asymptotics of the branching functions can be used to prove the finite reducibility theorem \cite{kac1988modular}. It states that for the conformal embedding  $\mathfrak{a}\subset\mathfrak{g}$ only finite number of branching coefficients have non-zero values.

Having found all such weights $\nu$ and the corresponding branching coefficients $b^{(\mu)}_{\nu}$ we can replace the characters of the $\mathfrak{g}$-modules in the diagonal modular-invariant partition function
\begin{equation}
  \label{eq:34}
   Z(\tau)=\sum_{ \mu\in P^{+}_{\mathfrak{g}}} \chi_{\mu}(\tau)\bar \chi_{\mu}(\bar \tau)
\end{equation}
by the decompositions
\begin{equation}
  \label{eq:32}
\sum_{\nu \in P^{+}_{\mathfrak{a}}}b^{(\mu)}_{\nu} \chi_{\nu}  
\end{equation}
 containing the modified characters $\chi_{\nu}$ of the corresponding $\mathfrak{a}$-modules. Thus we obtain the non-diagonal modular-invariant  partition function for the theory with the current algebra $\mathfrak{a}$.
\begin{equation}
  \label{eq:36}
   Z_{\mathfrak{a}}(\tau)=\sum_{ \nu,\lambda\in P^{+}_{\mathfrak{a}}} \chi_{\nu}(\tau)M_{\nu\lambda}\bar \chi_{\lambda}(\bar \tau)
\end{equation}

It is important to note that the orthogonal subalgebra $\mathfrak{a}_{\bot}$ is always empty for the conformal embeddings $\mathfrak{a}\subset \mathfrak{g}$. It can be seen from the following consideration. In the case of the conformal embedding energy-momentum tensors are to be equal
\begin{equation}
  \label{eq:46}
  T_{\mathfrak{a}}(z)=T_{\mathfrak{g}}(z)
\end{equation}
The energy-momentum tensor can be expanded into the modes $L_n$
\begin{equation}
\label{eq:47}
  T(z)=\frac{1}{2(k+h^v)}\sum_n z^{-n-1}L_n
\end{equation}
The modes are constructed as the combination of normally-ordered products of the generators of $\mathfrak{g}$.
\begin{equation}
\label{eq:48}
  L_n=\frac{1}{2(k+h^v)}\sum_a\sum_m:X^a_m X^a_{n-m}:
\end{equation}
The substitution of the generators of the algebra $\mathfrak{a}$  in terms of the generators of $\mathfrak{g}$ into these combinations of $T_{\mathfrak{a}}$ should give the energy-momentum tensor $T_{\mathfrak{g}}$. But if the set of the generators $\Delta_{\bot}$ is not empty it is not possible, since $T_{\mathfrak{g}}$ contains the combinations of the generators $X^{\alpha}_n, \; \alpha\in \Delta_{\bot}$, which can not be obtained in the expansion of components of $T_{\mathfrak{a}}$.

\subsection{Examples}
\label{sec:example}

@@

>>>>

\documentclass[a4paper,12pt]{article}
\usepackage{ucs}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{pb-diagram}
\usepackage{multicol}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian]{babel}
\usepackage{cmap}
\usepackage{color}
%\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\pagestyle{plain}
\usepackage[unicode,verbose,plainpages=false]{hyperref}

@@

\begin{document}
\pagenumbering{alph}
\thispagestyle{empty}

@@

\section{Finite dimensional Lie algebras}
\label{sec:finite-dimens-lie}

\subsection{Regular embedding of $A_1$ into $B_2$}
\label{sec:regul-embedd-a_1}

@@

\subsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the regular embedding $B_2 \longrightarrow B_4$.
We calculate the branching coefficients for the fundamental vector representation of $B_4$.
The corresponding Dynkin diagrams are in the Figure \ref{fig:dynkin}.
\begin{figure}[h]
  \centering
  \includegraphics[width=50mm]{B4_B2_2A1.pdf}
  \caption{Dynkin diagrams for the regular embedding $B_2\subset B_4$. Note the dropped node and the roots of $\mathfrak{a}_{\bot}$. }
  \label{fig:dynkin}
\end{figure}

@@

\begin{figure}[p]
  \centering
  \includegraphics[height=80mm]{B4_B2_fan.pdf}
  \caption{The fan for $B_2\subset B_4$}
  \label{fig:B4B2Fan}
\end{figure}
Applying the recurrent relation (\ref{recurrent-relation}) we obtain the
following branching coefficients:
\begin{multline}
  \label{eq:24}
  \pi_{\mathfrak{a}} \left(ch L^{(0,1,0,2)}_{B_4}\right) = 6 \; ch L^{(0,0)}_{B_2}+ 60
  \; ch L_{B_2}^{(0,2)}+ 30 \; ch L_{B_2}^{(1,0)}+ 19 \; ch L_{B_2}^{(2,0)}+\\
  40 \; ch L_{B_2}^{(1,2)}+ 10 \; ch L_{B_2}^{(2,2)}.
\end{multline}

\section{Applications to the conformal field theory}
\label{sec:phys-appl}

\subsection{Conformal embeddings}
\label{sec:conformal-embeddings}

Branching coefficients for an embedding of affine Lie subalgebra into
affine Lie algebra can be used to construct modular invariant
partition functions for Wess-Zumino-Novikov-Witten models of conformal field theory (\cite{difrancesco1997cft}, \cite{Walton:1999xc}, \cite{walton1989conformal}, \cite{schellekens1986conformal}). In these models current algebras are affine Lie algebras.

The modular invariant partition function is crucial for the conformal theory to be valid on the torus and higher genus Riemann surfaces. It is important for the applications of CFT to the string theory and to the critical phenomena description.

The simplest modular-invariant partition function has the diagonal form:
\begin{equation}
  \label{eq:34}
   Z(\tau)=\sum_{ \mu\in P^{+}_{\mathfrak{g}}} \chi_{\mu}(\tau)\bar \chi_{\mu}(\bar \tau)
\end{equation}
Here the sum is over the set of the highest weights of integrable modules of WZW-model and $\chi_{\mu}(\tau)$ are the modified characters of these modules.

The task of the construction of non-diagonal modular invariants is not easy. 
Consider the Wess-Zumino-Witten model with the affine Lie algebra $\mathfrak{a}$. Non-diagonal modular invariants for this model can be constructed from the diagonal invariant if there exists affine algebra $\mathfrak{g}$ such that $\mathfrak{a}\subset\mathfrak{g}$. 
Then we can replace the characters of the $\mathfrak{g}$-modules in the diagonal modular-invariant partition function (\ref{eq:36})
by the decompositions
\begin{equation}
  \label{eq:32}
\sum_{\nu \in P^{+}_{\mathfrak{a}}}b^{(\mu)}_{\nu} \chi_{\nu}  
\end{equation}
containing the modified characters $\chi_{\nu}$ of the corresponding $\mathfrak{a}$-modules. Thus we obtain the non-diagonal modular-invariant  partition function for the theory with the current algebra $\mathfrak{a}$.
\begin{equation}
  \label{eq:36}
   Z_{\mathfrak{a}}(\tau)=\sum_{ \nu,\lambda\in P^{+}_{\mathfrak{a}}} \chi_{\nu}(\tau)M_{\nu\lambda}\bar \chi_{\lambda}(\bar \tau)
\end{equation}
The reduction of the representations is crucial in this construction.

For the construction to be valid the embedding is required to preserve conformal invariance.
Let $X^{a_j}_{-n_j}$ and $\tilde{X}^{a'_j}_{-n_j}$ be the lowering generators for $\mathfrak{g}$ and for $\mathfrak{a}\subset\mathfrak{g}$ correspondingly. Let $\pi_{\mathfrak{a}}$ be the projection operator of $\pi_{\mathfrak{a}}:\mathfrak{g}\longrightarrow \mathfrak{a}$. In the theory attributed to $\mathfrak{g}$ with the vacuum $\left|\lambda\right>$ the states can be described as
\begin{equation}
  \label{eq:109}
  X^{a_1}_{-n_1}X^{a_2}_{-n_2}\dots\left|\lambda\right>\quad n_1\geq n_2\geq \dots>0.
\end{equation}
And for the sub-algebra $\mathfrak{a}$ the corresponding states are
\begin{equation}
  \label{eq:110}
  \tilde{X}^{a'_1}_{-n_1}\tilde{X}^{a'_2}_{-n_2}\dots\left|\pi_{\mathfrak{a}}(\lambda)\right>.
\end{equation}
The $\mathfrak{g}$-invariance of the vacuum entails its $\mathfrak{a}$-invariance, but this is not the case for the energy-momentum tensor. So the energy-momentum tensor of the larger theory should consist only of the generators of $\tilde{X}$. Then 
\begin{equation}
  \label{eq:2}
  T_{\mathfrak{g}}(z)=T_{\mathfrak{a}}(z),  
\end{equation}
which leads to the equality of the central charges 
\begin{equation}
  \label{eq:33}
  c(\mathfrak{g})=c(\mathfrak{a}).
\end{equation}
This leads to the equation
\begin{equation}
  \label{eq:111}
  \frac{k\;\mathrm{dim}\,\mathfrak{g}}{k+g}=\frac{x_e k\; \mathrm{dim}\,\mathfrak{a}}{x_ek+a}
\end{equation}
Here $x_e$ is the embedding index and $g$, $a$ are the dual Coxeter numbers for the  corresponding algebras.

@@

The relation (\ref{eq:111}) and the asymptotics of the branching functions can be used to prove the finite reducibility theorem \cite{kac1988modular}. It states that for the conformal embedding  $\mathfrak{a}\subset\mathfrak{g}$ only finite number of branching coefficients have non-zero values.

It is important to note that the orthogonal subalgebra $\mathfrak{a}_{\bot}$ is always empty for the conformal embeddings $\mathfrak{a}\subset \mathfrak{g}$. It can be seen from the following consideration. In the case of the conformal embedding energy-momentum tensors are to be equal (\ref{eq:2}).
The energy-momentum tensor can be expanded into the modes $L_n$
\begin{equation}
\label{eq:47}
  T(z)=\frac{1}{2(k+h^v)}\sum_n z^{-n-1}L_n
\end{equation}
The modes are constructed as the combination of normally-ordered products of the generators of $\mathfrak{g}$.
\begin{equation}
\label{eq:48}
  L_n=\frac{1}{2(k+h^v)}\sum_a\sum_m:X^a_m X^a_{n-m}:
\end{equation}
The substitution of the generators of the algebra $\mathfrak{a}$  in terms of the generators of $\mathfrak{g}$ into these combinations of $T_{\mathfrak{a}}$ should give the energy-momentum tensor $T_{\mathfrak{g}}$. But if the set of the generators $\Delta_{\bot}$ is not empty it is not possible, since $T_{\mathfrak{g}}$ contains the combinations of the generators $X^{\alpha}_n, \; \alpha\in \Delta_{\bot}$, which can not be obtained in the expansion of components of $T_{\mathfrak{a}}$.

@@

\subsection{Coset models}
\label{sec:coset-models}
